### 2017-08-05 KineticModel/estimation/docs/formulation/ident.tex
no major change
add all estimation formulations

	\subsubsection{Identifiability analysis of parameters in a kinetic model of gluconeogenesis:}
	The proposed model for acetate consumption through gluconeogenesis and its corresponding kinetic model is used as a case study to illustrate the utility of identifiability analysis for the design of experiments for estimating parameters in kinetic models of metabolism. The kinetic model is described below.
	\begin{equation}\label{eq:ode1}
	\frac{d}{dt}pep=v_1-v_2-v_4
	\end{equation}
	\begin{equation}\label{eq:ode2}
	\frac{d}{dt}fdp=v_2-v_3
	\end{equation}
	\begin{equation}\label{eq:ode3}
	\frac{d}{dt}E=v_{e,max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right) - d E
	\end{equation}
	The kinetic expressions for fluxes $v_1$ through $v_4$ are given below. The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics. The acetate flux through $v_1$ is also governed by the quantity of available enzyme E. 
	\begin{equation}\label{eq:flux1}
	v_1 = k_{1}^{cat}E\frac{acetate}{acetate+K_{1}^{acetate}}
	\end{equation}	
	\begin{equation}\label{eq:flux2}
	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
	\end{equation}
	\begin{equation}\label{eq:flux3}
	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
	\end{equation}
	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$. The added flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
	\begin{equation}\label{eq:flux4}
	v_4 = k_{4}^{cat}.pep
	\end{equation}
\end{document}

>>>>

	\subsubsection{Identifiability analysis of parameters in a kinetic model of gluconeogenesis:}
	The proposed model for acetate consumption through gluconeogenesis and its corresponding kinetic model is used as a case study to illustrate the utility of identifiability analysis for the design of experiments for estimating parameters in kinetic models of metabolism. The kinetic model is described below.
	\begin{equation}\label{eq:ode1}
	\frac{d}{dt}pep=v_1-v_2-v_4
	\end{equation}
	\begin{equation}\label{eq:ode2}
	\frac{d}{dt}fdp=v_2-v_3
	\end{equation}
	\begin{equation}\label{eq:ode3}
	\frac{d}{dt}E=v_{e,max}\left(\frac{1}{1+\left(\frac{fdp}{K_{e}^{fdp}}\right)^{n_e}}\right) - d E
	\end{equation}
	The kinetic expressions for fluxes $v_1$ through $v_4$ are given below. The consumption of acetate through $v_1$ and conversion of \textit{pep} through $v_2$ are expressed in Equations (\ref{eq:flux1}) and (\ref{eq:flux2}) respectively using Michaelis-Menten kinetics. The acetate flux through $v_1$ is also governed by the quantity of available enzyme E. 
	\begin{equation}\label{eq:flux1}
	v_1 = k_{1}^{cat}E\frac{acetate}{acetate+K_{1}^{acetate}}
	\end{equation}	
	\begin{equation}\label{eq:flux2}
	v_2 = V_{2}^{max}\frac{pep}{pep+K_{2}^{pep}}
	\end{equation}
	\begin{equation}\label{eq:flux3}
	v_3 = V_{3}^{max}\frac{\tilde{fdp}\left(1+\tilde{fdp}\right)^3}{\left(1+\tilde{fdp}\right)^4+L_3\left(1+\frac{pep}{K_{3}^{pep}}\right)^{-4}}
	\end{equation}
	The allosterically regulated flux $v_3$ for the consumption of \textit{fdp} is expressed in Equation (\ref{eq:flux3}) using the Monod-Wyman-Changeux (MWC) model for allosterically regulated enzymes, where $\tilde{fdp}$ refers to the ratio of \textit{fdp} with respect to its allosteric binding constant $K_{3}^{fdp}$. The added flux $v_4$ for the export of \textit{pep} is expressed as a linear equation dependent on $pep$ in Equation (\ref{eq:flux4}).
	\begin{equation}\label{eq:flux4}
	v_4 = k_{4}^{cat}.pep
	\end{equation}
	Simulated data

	\subsection{Parameter estimation in kinetic models of metabolism:}
	The formulation for parameter estimation assumes that the concentrations and the fluxes corresponding to each perturbation are variables in the optimization problem formulated to estimate kinetic parameters.

	Let $\mathbf{x}\in\mathbb{R}^m$, $\mathbf{p}\in\mathbb{R}^l$ and $\mathbf{v}\in\mathbb{R}$ be the vector of concentrations, parameters and fluxes respectively. If the number of perturbations used for parameter estimation is in $\mathbb{R}^p$, then the vector of concentrations, fluxes and parameters that need to be estimated in the optimization problem changes to $\mathbf{x}\in\mathbb{R}^{mp}, \mathbf{v}\in\mathbb{R}^{p}$ and $\mathbf{p}\in\mathbb{R}^{l}$ respectively.

	Let $i\in\mathbb{R}^{mp}$, $j\in\mathbb{R}^{p}$ and $k\in\mathbb{R}^{l}$ represent the indices for concentrations, fluxes and parameters respectively in Equation (\ref{eq:l2norm-flux}).

	\begin{center}
		\begin{subequations}\label{eq:l2norm-flux}
			\begin{align}
			\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert\\
			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
			& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
			& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
			\end{align}
		\end{subequations}
	\end{center}

	Due to the inability of the above formulation to get feasible results, we modified the problem to include the L2-norm of the concentrations as well in the objective function in Equation (\ref{eq:free-var}a). In Equation (\ref{eq:free-var}) the flux and the concentrations are free variables, i.e.,  $\mathbf{x}\in\left[\mathbf{0},+\mathbf{\infty}\right]$. 
	\begin{center}
		\begin{subequations}\label{eq:free-var}
			\begin{align}
			\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
			& \text{      }0\le x_i \le +\infty & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
			& \text{      }0\le v_j \le +\infty & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
			\end{align}
		\end{subequations}
	\end{center}
After testing the problem in Equation (\ref{eq:free-var}) above for the acetate uptake flux $v_1$, we found that all solutions using IPOPT exceeded the maximum number of iterations (set at 5000) and never produced an optimal solution.

	We further also test a formulation in Equation (\ref{eq:l2norm-conc}) where the flux and concentrations are constrained as earlier with a noise term.	
	\begin{center}
		\begin{subequations}\label{eq:l2norm-conc}
			\begin{align}
			\underset{\mathbf{x},\mathbf{p},\mathbf{v}}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert\\
			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
			& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
			& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
			\end{align}
		\end{subequations}
	\end{center}

	A third formulation where the noise terms ($\epsilon$) used as bounds for the flux and the concentrations are included in the optimization problem as a variable to be penalized is given in Equation (\ref{eq:noise-var}) below.

	\begin{center}
		\begin{subequations}\label{eq:noise-var}
			\begin{align}
			\underset{\mathbf{x},\mathbf{p},\mathbf{v},\epsilon}{\mathrm{min}} & \text{      }w_1\left\Vert \mathbf{v}-\mathbf{v}^*\right\Vert + w_2\left\Vert \mathbf{x}-\mathbf{x}^*\right\Vert + w_3\epsilon^2\\
			\mathrm{st}& \text{      }N(\mathbf{x},\mathbf{p}) - v_jD(\mathbf{x},\mathbf{p}) = 0 & \text{  }\forall & \text{ }j\in\mathbb{R}^p\\
			& \text{      }x^*_i(1-\epsilon)\le x_i \le x^*_i(1+\epsilon) & \text{  }\forall & \text{ }i\in\mathbb{R}^{mp}\\	
			& \text{      }v^*_j(1-\epsilon)\le v_j \le v^*_j(1+\epsilon) & \text{  }\forall & \text{ }j\in\mathbb{R}^{p}\\		
			& \text{      }\mathbf{p}_{min} \le \mathbf{p} \le \mathbf{p}_{max}
			\end{align}
		\end{subequations}
	\end{center}

\end{document}
